Modelling $q_{95}$ Windows for the Suppression of Edge Localized Modes by Resonant Magnetic Perturbations in the DIII-D Tokamak

### 2020-07-05 ExperimentalData/DIIID/145380/Paper/paper.tex
First draft of 145380 paper

\documentclass[12pt,prb,aps]{revtex4-1}
\usepackage{amsmath}
\pdfoutput=1 
\usepackage{graphicx}
\newcommand {\bomega}{\mbox{\boldmath$\omega$}}
\newcommand {\bpi}{\mbox{\boldmath$\pi$}}

@@

\author{R.~Fitzpatrick}
\affiliation{Institute for Fusion Studies,  Department of Physics,  University of Texas at Austin,  Austin TX, 78712, USA}

@@

\section{Introduction}
Tokamak discharges operating in high-confinement mode (H-mode)\,\cite{wagner} exhibit intermittent bursts of heat and particle transport, 
emanating from the outer regions of the plasma, that are known as ``type-I edge localized modes'' (ELMs). \cite{zohm}
It is estimated that the heat load that ELMs
will deliver to the tungsten plasma-facing components in a reactor-scale tokamak, such as ITER, will be large enough to cause
massive tungsten ion influx into the plasma core, and that the erosion associated with this process will 
unacceptably limit the lifetimes of these components. \cite{loarte} Consequently, the development of robust and effective
methods for ELM control is a high priority for the international magnetic fusion program.

@@

Recently, a toroidal generalization of the 
cylindrical asymptotic matching model presented
in Ref.~\onlinecite{rf1} was formulated and
used to model  RMP-induced ELM-suppression experiments performed on the DIII-D tokamak, \cite{rf2}
leading to similar conclusions to the aforementioned TM1
studies. The primary aim of this paper is to employ this
new model to try to account for the  $q_{95}$
ELM-suppression windows  that are apparent when the edge safety-factor is slowly ramped in
a particular DIII-D discharge in which an $n=3$
RMP is used to control ELMs.\cite{d3d,d3d2}

It is well-known that magnetic reconnection driven when a  (stable) tearing mode interacts with a static RMP that is resonant at a particular magnetic flux-surface in a tokamak plasma is
facilitated when the associated ``natural frequency'' is relatively small.\cite{rfa,rfb}  The natural frequency  of a  (stable) tearing mode is the
helical phase velocity that the mode would possess were it naturally unstable (in the absence of the RMP).  Driven magnetic reconnection leads to the
formation of a locked magnetic island chain at the resonant surface in question. 
Hence, the prevailing hypothesis is that a $q_{95}$ window for
RMP-induced ELM suppression occurs when $q_{95}$ is such that the natural frequency of a tearing mode resonant  at 
the top of the pedestal is close to zero. 
However, there is currently some uncertainty in the fusion community regarding the form of the natural frequency.

According to
linear tearing mode theory, a tearing mode is essentially convected by
the local electron fluid at the resonant surface (see the discussion in Sect.~\ref{natural}).\cite{hender,cole} Hence, if the response of an H-mode tokamak plasma to an applied RMP
is governed by linear physics then we would expect the
natural frequency to take the form\,\cite{lin1,lin2,lin3}
\begin{equation}
\varpi_{\rm linear} = - n\,(\omega_E+\omega_{\ast\,e}),\label{e1}
\end{equation}
where
\begin{align}
\omega_E &= -\frac{d{\mit\Phi}}{d{\mit\Psi}_p},\\[0.5ex]
\omega_{\ast\,e} &=\frac{T_e}{e}\,\frac{d\ln p_e}{d{\mit\Psi}_p}.
\end{align}
Here, ${\mit\Psi}_p$ is the equilibrium poloidal magnetic
flux (divided by $2\pi$), ${\mit\Phi}({\mit\Psi}_p)$ the equilibrium scalar
electric potential, $p_e({\mit\Psi}_p)$ the equilibrium
electron pressure, $T_e({\mit\Psi}_p)$ the equilibrium
electron temperature, $e$ the magnitude of the electron
charge, and $n$ the toroidal mode number of the RMP. 
Moreover, the right-hand side of Eq.~(\ref{e1}) is evaluated
at the ``rational''  (i.e., resonant) magnetic flux-surface at which the
safety-factor
\begin{equation}
q({\mit\Psi}_p)= \frac{d{\mit\Psi}_t}{d{\mit\Psi}_p},
\end{equation}
takes the rational value $n/m$, where $m$ is a positive integer. Here, $m$ and $n$ are the numbers of poloidal and
toroidal periods, respectively, of the helical magnetic island chain driven at the rational surface.  Furthermore, ${\mit\Psi}_t({\mit\Psi}_p)$ is the equilibrium toroidal magnetic flux (divided by $2\pi$).

According to nonlinear tearing mode theory, a tearing mode is essentially convected by
the local ion fluid at the rational surface (again, see the discussion in Sect.~\ref{natural}).\cite{nl1,nl2,nl3}
In fact, if the response of an H-mode tokamak plasma to an applied RMP
is governed by nonlinear physics then we would expect the
natural frequency to take the form\,\cite{rf2}
\begin{equation}\label{e2}
\varpi_{\rm nonlinear} = -n\left(\omega_E +\left[1-L_{00}^{\,ii}+L_{01}^{\,ii}\left(\frac{\eta_i}{1+\eta_i}\right)\right]
\omega_{\ast\,i}
%\right.\nonumber\\[0.5ex]&\left.
-\left[L_{00}^{\,iI}- L_{01}^{\,iI}\left(\frac{\eta_I}{1+\eta_I}\right)\right]\omega_{\ast\,I}\right),
\end{equation}
where
\begin{align}
\omega_{\ast\,a}({\mit\Psi}_p) &=-\frac{T_a}{Z_a\,e}\,\frac{d\ln p_a}{d{\mit\Psi}_p},\\[0.5ex]
\eta_a({\mit\Psi}_p)&=\frac{d\ln T_a}{d\ln n_a},
\end{align}
for $a=i, I$. 
Here,  $Z_i$, $n_i({\mit\Psi}_p)$, $T_i({\mit\Psi}_p)$, and $p_i({\mit\Psi}_p)=n_i\,T_i$   are the charge number, equilibrium number density, 
equilibrium temperature, and equilibrium pressure
of the majority (thermal) ions, respectively,
whereas  $Z_I$, $n_I$, $T_I$, $p_I = n_I\,T_I$
are the corresponding quantities for the impurity ions.  
Furthermore, $L_{00}^{ii}({\mit\Psi}_p)$, $L_{01}^{ii}({\mit\Psi}_p)$,  $L_{00}^{II}({\mit\Psi}_p)$, and $L_{01}^{II}({\mit\Psi}_p)$ are neoclassical parameters that are defined
in Ref.~\onlinecite{rf2}. Note that these parameters are
affected by charge exchange  with neutrals. Finally,
as before, the right-hand side of Eq.~(\ref{e2}) is evaluated
at the rational magnetic flux-surface.

A third possibility is that a tearing mode is convected by the local guiding center fluid at the rational surface,\cite{heyn,paz1}
in which case we would expect the natural frequency to take the form 
\begin{equation}\label{e3}
\varpi_{EB} =-n\,\omega_E.
\end{equation}
As before, the right-hand side of Eq.~(\ref{e3}) is evaluated
at the rational magnetic flux-surface.

@@

\section{Description of Theoretical Model}
The theoretical model of the response of a tokamak plasma to an externally applied RMP that is used in this
paper is described in detail in Ref.~\onlinecite{rf2}. The model employs a standard asymptotic matching approach.\cite{fkr,am1,am2,am3}
According to this approach, the response of the plasma to the applied RMP is governed by a combination of flux-freezing and
perturbed force balance (this combination is often referred to as ``marginally-stable ideal-MHD", which is a misnomer because MHD {\em per se}\/ plays no
role) everywhere in the plasma apart from a number of relatively narrow (in the radial direction) regions in which the applied
perturbation resonates with the equilibrium magnetic field. Magnetic reconnection can take place within the resonant regions to
produce relatively thin magnetic islands. Within the resonant regions, the plasma response is governed by nonlinear,
as opposed to linear, two-fluid resistive MHD. This is the case because the widths of the magnetic island chains
driven at the resonant surfaces exceed the linear layer widths (which invalidates linear theory).\cite{rf1}
Thus, when employing the asymptotic matching approach, the equations of flux-freezing and perturbed force balance
are solved in the so-called ``outer region'' that comprises most of the plasma (and the surrounding vacuum), the equations of 
nonlinear two-fluid resistive-MHD are solved in the various resonant layers that constitute the so-called ``inner region'', and the two
sets of solutions are then asymptotically matched to one another.

A toroidal tokamak equilibrium exhibits two distinct types of response to an applied RMP.\cite{paz1,kink0,kink1}
The first of these is known as the ``tearing response''---this is a non-ideal-MHD response that is associated with the
formation of current sheets and magnetic island chains at various resonant surfaces within the plasma. The second
response type is known as the ``kink response''---this is an edge-localized ideal-MHD response that is associated
with coupling to a stable non-resonant kink mode. For the case of the tearing response, our model employs an
approximation in which the plasma response is assumed to be vacuum-like between the various resonant surfaces. On the other hand,
the kink response of the plasma is calculated exactly using the {\tt GPEC} code.\cite{ipec,gpec}

Our model has been implemented in the {\tt EPEC} (Extended Perturbed Equilibrium Code) code. The name
of this code reflects the fact that the nonlinear evolution of tearing modes in a
tokamak plasma has far more in common with the $1\tfrac{1}{2}$-D
evolution of the global plasma equilibrium than it does with
conventional linear tearing mode physics. (See the discussion in Sect.~\ref{general}.) In particular, the Alfv\'{e}n time is
an irrelevant timescale in nonlinear tearing mode theory, and
is also  very much shorter than the timescales on which
physical quantities of interest actually evolve. (Note that all timescales are normalized to the Alfv\'{e}n time
in Ref.~\onlinecite{rf2}. However, this is just a matter of convention. With the benefit of hindsight, it would
have been better to normalize the timescales with respect to a diamagnetic timescale, in which case the Alfv\'{e}n time would have completely dropped
out of the final system of equations.)

\section{Modeling of DIII-D Discharge \#145380}
\subsection{Plasma Equilibrium}
DIII-D discharge \#145380 is an ITER Similar Shape (ISS), ELMing, H-mode discharge with a toroidal magnetic field $B_T=-1.9\,{\rm T}$ in which the
plasma current $I_p$ is slowly ramped over a 2 s interval to scan the magnetic safety-factor.\cite{d3d,d3d2} The majority ions are Deuterium, whereas the
minority ions are Carbon~VI.
Figure~\ref{fig1} gives an overview of this discharge. Note that the pedestal density is relatively high ($n_{e\,{\rm ped}}\sim 3\times 10^{19}\,{\rm m}^{-3}$). 
A static RMP is applied to the plasma by running steady $n=3$ currents through the I-coil system.\cite{icoil}
Three windows of ELM suppression are evident: the first extends from $t=2840$--$2980$ ms; the
second extends from $t=3320$--$3560$ ms; and the third extends from $t=3880$--$4200$ ms.

Figure~\ref{fig2} shows the measured plasma equilibrium at the start of the current ramp ($t=2500$ ms). Likewise,
Fig.~\ref{fig3} shows the measured plasma profiles at the start of the current ramp. Here, ${\mit\Psi}_N$ is a
normalized poloidal magnetic flux defined such that the magnetic axis corresponds to ${\mit\Psi}_N=0$, 
and the last closed magnetic flux-surface (LCFS) to ${\mit\Psi}_N=1$. The perpendicular
diffusivity data comes from {\tt TRANSP} modelling.\cite{transp}

Measured plasma equilibrium and
plasma profile data is available every $50$ ms. The equilibrium and profile data at a given point in time is interpolated from
the measured data. In fact, quantities that depend on the plasma equilibrium and profiles are recalculated every 1 ms.  Data from the
{\tt GPEC} code is available every 100 ms. As before, the {\tt GPEC} data at a given point in time is interpolated from the
available {\tt GPEC} data, and any quantity that depends on this data is recalculated every 1 ms. There is one major exception to
the aforementioned scheme. The theoretical model used in this paper actually requires plasma equilibria and profiles that are unaffected
by the applied RMP. However, there is clear evidence that the equilibria and the profiles (in particular) are significantly modified
within the three ELM-suppression windows (presumably because wide magnetic island chains are driven in the pedestal). 
Hence, we do not use the equilibrium, profile, or {\tt GPEC}, data within the three ELM-suppression windows. Instead, we
interpolate across these windows using data taken on either side of the windows.

Figure~\ref{fig4} shows the safety-factor in DIII-D discharge \#145380 as a function of time. It can be seen
that $q_{95}\equiv q({\mit\Psi}_N=0.95)$ varies from about 4.1 at the start of the current ramp to about 3.3
at the end. However, the safety-factor at the magnetic axis, $q_0 \equiv q({\mit\Psi}_N=0.0)$, remains approximately
constant at about 1.0. For the sake of practicality,  the {\tt GPEC} and {\tt EPEC} calculations described in this paper ignore rational
surfaces that lie very close to the LCFS: in other words,  surfaces that lie in the region $0.997 < {\mit\Psi}_N\leq 1.000$.

Note that every quantity used in our theoretical model is derived from experimental measurements taken in DIII-D discharge \#145380, with one
exception. The exception is the neutral particle data. Unfortunately, no neutral particle data is available for DIII-D discharge \#145380.
Hence, we employ neutral particle data derived from a much earlier discharge (see Ref.~\onlinecite{rf2}).\cite{neutral}
In future, data from the new {\tt LLAMA} (Lyman Alpha Measurement Apparatus) system will be available in DIII-D discharges.\cite{llama}
Combining this data  with modeling from the SOLPS-ITER code\,\cite{solps} ought to generate sufficient information to fully determine the
neutral particle parameters required by the {\tt EPEC} code.

\subsection{Case 1: $\varpi=\varpi_{\rm nonlinear}$}
This first case that we shall consider is that in which the natural frequency is determined by nonlinear
tearing mode theory: in other words, the natural frequency is specified by Eq.~(\ref{e2}).

The upper panel of Fig.~\ref{fig5} shows the natural frequencies of the various  $n=3$ tearing modes 
 that are resonant in the pedestal. It can be seen that the natural frequency of the $m=11$ harmonic has its main zero crossing in the middle
of the first ELM-suppression window. Likewise, the natural frequency of the $m=10$ harmonic has its main zero crossing in the middle of the
second ELM-suppression window. However, the natural frequency of the $m=9$ harmonic has its main zero crossing slightly
before the third ELM-suppression window. The lower panel of Fig.~\ref{fig5} shows the actual phase velocities of the various $n=3$
tearing modes that are resonant in the pedestal. It can be seen that if the magnitude of the natural frequency falls below about
$5\,{\rm krad/s}$ then the associated tearing mode locks to the RMP (i.e., its phase velocity becomes zero).

\subsection{Case 2: $\varpi=\varpi_{EB}$}

\subsection{Case 3: $\varpi=\varpi_{\rm linear}$}

\section{Summary and Discussion}

>>>>

\documentclass[12pt,prb,aps]{revtex4-1}
\usepackage {amsmath}
\pdfoutput = 1 
\usepackage {graphicx}
\newcommand {\bomega}{\mbox{\boldmath$\omega$}}
\newcommand {\bpi}{\mbox{\boldmath$\pi$}}

@@

\author{R.~Fitzpatrick\,\footnote{rfitzp@farside.ph.utexas.edu}}
\affiliation{Institute for Fusion Studies,  Department of Physics,  University of Texas at Austin,  Austin TX, 78712, USA}

\begin{abstract}
A toroidal asymptotic matching model of the response of a tokamak plasma to a static resonant magnetic perturbation (RMP) is used to simulate the $n=3$ RMP-induced 
edge-localized-mode (ELM)-suppression
windows in $q_{95}$ that are evident when the plasma current is slowly ramped in DIII-D discharge \#145380. All quantities employed in the simulation are
derived from experimental measurements, apart from the neutral particle data. 
Three cases are considered.
In the first case, the natural frequencies of tearing modes resonant in the plasma are 
determined by the ion flows at the corresponding rational surfaces, which is the prediction of nonlinear theory.
In the second case, the natural frequencies are 
determined by the guiding center flows at the rational surfaces.
 In the third case, the natural frequencies   are 
determined by the electron flows at the  rational surfaces, which is the prediction of linear theory.
The second case gives  the best agreement between the simulations and the  experimental observations.   The third case gives the worst agreement between the simulations 
and the  observations. The first case only leads to partial agreement between  the  simulations and the  observations. However, this lack of complete agreement
may be a consequence of using an inaccurate guess for the neutral particle distribution in the pedestal.  
\end{abstract}

@@

\section{Introduction}
Tokamak discharges operating in high-confinement mode (H-mode)\,\cite{wagner} exhibit intermittent bursts of heat and particle transport, 
emanating from the outer regions of the plasma, that are known as  (type-I) ``edge localized modes'' (ELMs).\cite{zohm}
It is estimated that the heat load that ELMs
will deliver to the tungsten plasma-facing components in a reactor-scale tokamak, such as ITER, will be large enough to cause
massive tungsten ion influx into the plasma core, and that the erosion associated with this process will 
unacceptably limit the lifetimes of these components.\cite{loarte} Consequently, the development of robust and effective
methods for ELM control is a high priority for the international magnetic fusion program.

@@

Recently, a toroidal generalization of the 
cylindrical asymptotic matching model presented
in Ref.~\onlinecite{rf1} was formulated and
used to model  RMP-induced ELM-suppression experiments performed on the DIII-D tokamak,\cite{rf2}
leading to similar conclusions to the aforementioned TM1
studies. The primary aim of this paper is to employ this
new model to try to account for the  $q_{95}$
ELM-suppression windows  that are apparent when the edge safety-factor is slowly ramped in
a particular DIII-D discharge (\#145380) in which an $n=3$
RMP is used to control ELMs.\cite{d3d,d3d2}

It is well-known that magnetic reconnection driven when a  (stable) tearing mode interacts with a static RMP that is resonant at a particular magnetic flux-surface in a tokamak plasma is
facilitated when the associated ``natural frequency'' is relatively small.\cite{rfa,rfb}  The natural frequency  of a  (stable) tearing mode is the
helical phase velocity that the mode would possess were it naturally unstable (in the absence of the RMP).  Driven magnetic reconnection leads to the
formation of a locked magnetic island chain at the resonant surface in question. 
Hence, the prevailing hypothesis is that a $q_{95}$ window for
RMP-induced ELM suppression occurs when $q_{95}$ is such that the natural frequency of a tearing mode resonant  at 
the top of the pedestal is close to zero.\cite{d3d} 
However, there is currently some uncertainty in the fusion community regarding the form of the natural frequency.

According to nonlinear tearing mode theory, a tearing mode is essentially convected by
the local ion fluid at the rational surface (see the discussion in Sect.~\ref{natural}).\cite{nl1,nl2,nl3}
In fact, if the response of an H-mode tokamak plasma to an applied RMP
is governed by nonlinear physics then we would expect the
natural frequency to take the form\,\cite{rf2}
\begin{equation}\label{e2}
\varpi_{\perp\,i} = -n\left(\omega_E +\left[1-L_{00}^{\,ii}+L_{01}^{\,ii}\left(\frac{\eta_i}{1+\eta_i}\right)\right]
\omega_{\ast\,i}
-\left[L_{00}^{\,iI}- L_{01}^{\,iI}\left(\frac{\eta_I}{1+\eta_I}\right)\right]\omega_{\ast\,I}\right),
\end{equation}
where
\begin{align}
\omega_E({\mit\Psi}_p) &= -\frac{d{\mit\Phi}}{d{\mit\Psi}_p},\\[0.5ex]
\omega_{\ast\,a}({\mit\Psi}_p) &=-\frac{T_a}{Z_a\,e}\,\frac{d\ln p_a}{d{\mit\Psi}_p},\\[0.5ex]
\eta_a({\mit\Psi}_p)&=\frac{d\ln T_a}{d\ln n_a},
\end{align}
for $a=i, I$. 
Here,  $e$ is the magnitude of the electron charge, ${\mit\Psi}_p$  the equilibrium poloidal magnetic
flux (divided by $2\pi$), ${\mit\Phi}({\mit\Psi}_p)$ the equilibrium scalar
electric potential, and $n$ the toroidal mode number of the RMP.  Moreover, $Z_i$, $n_i({\mit\Psi}_p)$, $T_i({\mit\Psi}_p)$, and $p_i({\mit\Psi}_p)=n_i\,T_i$   are the charge number, 
equilibrium number density, 
equilibrium temperature, and equilibrium pressure
of the majority (thermal) ions, respectively,
whereas  $Z_I$, $n_I$, $T_I$, $p_I = n_I\,T_I$
are the corresponding quantities for the impurity ions.  
Furthermore, $L_{00}^{ii}({\mit\Psi}_p)$, $L_{01}^{ii}({\mit\Psi}_p)$,  $L_{00}^{II}({\mit\Psi}_p)$, and $L_{01}^{II}({\mit\Psi}_p)$ are neoclassical parameters that are defined
in Sect.~B of Ref.~\onlinecite{rf2}. Note that these parameters are
affected by charge exchange  with neutrals. 
The right-hand side of Eq.~(\ref{e2}) is evaluated
at the ``rational''  (i.e., resonant) magnetic flux-surface at which the
safety-factor
\begin{equation}
q({\mit\Psi}_p)= \frac{d{\mit\Psi}_t}{d{\mit\Psi}_p},
\end{equation}
takes the rational value $n/m$, where $m$ is a positive integer. Here, $m$ and $n$ are the numbers of poloidal and
toroidal periods, respectively, of the helical magnetic island chain driven at the rational surface. Moreover, ${\mit\Psi}_t({\mit\Psi}_p)$ is the equilibrium toroidal magnetic flux (divided by $2\pi$).

A second possibility is that a tearing mode is convected by the local guiding center fluid at the rational surface,\cite{heyn,paz1}
in which case we would expect the natural frequency to take the form 
\begin{equation}\label{e3}
\varpi_{\perp\, EB} =-n\,\omega_E.
\end{equation}
As before, the right-hand side of Eq.~(\ref{e3}) is evaluated
at the rational magnetic flux-surface.

Finally, according to
linear tearing mode theory, a tearing mode is essentially convected by
the local electron fluid at the resonant surface (again, see the discussion in Sect.~\ref{natural}).\cite{hender,cole} Hence, if the response of an H-mode tokamak plasma to an applied RMP
is governed by linear physics then we would expect the
natural frequency to take the form\,\cite{lin1,lin2,lin3}
\begin{equation}
\varpi_{\perp\,e} = - n\,(\omega_E+\omega_{\ast\,e}),\label{e1}
\end{equation}
where
\begin{align}
\omega_{\ast\,e}({\mit\Psi}_p) &=\frac{T_e}{e}\,\frac{d\ln p_e}{d{\mit\Psi}_p}.
\end{align}
 Here, $p_e({\mit\Psi}_p)$ is the equilibrium
electron pressure, and $T_e({\mit\Psi}_p)$ the equilibrium
electron temperature.
Finally,
as before, the right-hand side of Eq.~(\ref{e1}) is evaluated
at the rational magnetic flux-surface.

@@

\section{Description of Theoretical Model}
The theoretical model of the response of a tokamak plasma to an externally applied RMP that is used in this
paper is described in detail in Ref.~\onlinecite{rf2}. The model employs a standard asymptotic matching approach.\cite{fkr,am1,am2,am3}
According to this approach, the response of the plasma to the applied RMP is governed by a combination of flux-freezing and
perturbed force balance (this combination is often referred to as ``marginally-stable ideal-MHD", which is a misnomer because MHD {\em per se}\/ plays no
role) everywhere in the plasma apart from a number of relatively narrow (in the radial direction) regions in which the applied
perturbation resonates with the equilibrium magnetic field. Magnetic reconnection can take place within the resonant regions to
produce relatively thin magnetic islands. Within the resonant regions, the plasma response is governed by nonlinear,
as opposed to linear, two-fluid resistive-MHD. This is the case because the widths of the magnetic island chains
driven at the resonant surfaces exceed the linear layer widths (which invalidates linear theory).\cite{rf1}
Thus, when employing the asymptotic matching approach, the equations of flux-freezing and perturbed force balance
are solved in the so-called ``outer region'' that comprises most of the plasma (and the surrounding vacuum), the equations of 
nonlinear two-fluid resistive-MHD are solved in the various resonant layers that constitute the so-called ``inner region'', and the two
sets of solutions are then asymptotically matched to one another.

A toroidal tokamak equilibrium exhibits two distinct types of response to an applied RMP.\cite{paz1,kink0,kink1}
The first of these is known as the ``tearing response''---this is a non-ideal-MHD response that is associated with the
formation of current sheets and magnetic island chains at various resonant surfaces within the plasma. The second
response type is known as the ``kink response''---this is an edge-localized ideal-MHD response that is associated
with coupling to a stable non-resonant kink mode. For the case of the tearing response, our model employs an
approximation in which the plasma response is assumed to be vacuum-like between the various resonant surfaces. On the other hand,
the kink response of the plasma is calculated exactly using the {\tt GPEC} code.\cite{ipec,gpec} (See Sect.~A.8 of Ref.~\onlinecite{rf2}.)

Our model has been implemented in the {\tt EPEC} (Extended Perturbed Equilibrium Code) code. The name
of this code reflects the fact that the nonlinear evolution of tearing modes in a
tokamak plasma has far more in common with the $1\tfrac{1}{2}$-D
evolution of the global plasma equilibrium than it does with
conventional linear tearing mode physics. (See the discussion in Sect.~\ref{general}.) In particular, the Alfv\'{e}n time is
an irrelevant timescale in nonlinear tearing mode theory, and
is also  very much shorter than the timescales on which
physical quantities of interest actually evolve. (Note that all timescales are normalized to the Alfv\'{e}n time
in Ref.~\onlinecite{rf2}. However, this is just a matter of convention. With the benefit of hindsight, it would
have been better to normalize the timescales with respect to a diamagnetic timescale, in which case the Alfv\'{e}n time would have completely dropped
out of the final system of equations. See Sect.~E in Ref.~\onlinecite{rf2}.)

\section{EPEC Modeling of DIII-D Discharge \#145380}
\subsection{Plasma Equilibrium}
DIII-D discharge \#145380 is an ITER-Similar-Shape (ISS), ELMing, H-mode discharge, with a toroidal magnetic field $B_T=-1.9\,{\rm T}$, in which the
plasma current, $I_p$, is slowly ramped over a 2 second interval in order to scan the magnetic safety-factor.\cite{d3d,d3d2} The majority ions are Deuterium, whereas the
minority ions are Carbon~VI.
Figure~\ref{fig1} gives an overview of DIII-D discharge \#145380. Note that the pedestal electron number density is relatively high ($n_{e\,{\rm ped}}\sim 3\times 10^{19}\,{\rm m}^{-3}$). 
A static RMP is applied to the plasma by running steady $n=3$ currents through the I-coil system.\cite{icoil}
Three windows of ELM suppression are evident: the first extends from $t=2840$--$2980$ ms; the
second extends from $t=3320$--$3560$ ms; and the third extends from $t=3880$--$4200$ ms.

Figure~\ref{fig2} shows the measured plasma equilibrium at the start of the current ramp ($t=2500$ ms). Likewise,
Fig.~\ref{fig3} shows the measured plasma profiles at the start of the current ramp. Here, ${\mit\Psi}_N$ is a
normalized equilibrium poloidal magnetic flux defined such that the magnetic axis corresponds to ${\mit\Psi}_N=0$, 
and the last closed magnetic flux-surface (LCFS) to ${\mit\Psi}_N=1$. The perpendicular
diffusivity data comes from {\tt TRANSP} modelling.\cite{transp}

Measured plasma equilibrium and
plasma profiles are available every $50$ ms. The equilibrium and profiles  at a given point in time are interpolated from
the measured equilibria and profiles. In fact, all quantities that depend on the plasma equilibrium and profiles are recalculated every 1 ms.  Data values from the
{\tt GPEC} code are available every 100 ms. As before, the {\tt GPEC} data values at a given point in time are interpolated from the
available {\tt GPEC} data values, and any quantity that depends on these values is recalculated every 1 ms. There is one major exception to
the aforementioned scheme. The theoretical model used in this paper actually requires plasma equilibria and profiles that are unaffected
by the applied RMP. However, there is clear evidence that the equilibria and the profiles (in particular) are significantly modified
within the three ELM-suppression windows (presumably because wide magnetic island chains are driven in the pedestal). [See Fig.~1(c).]
Hence, we do not use the equilibrium, profile, or {\tt GPEC}, data within the three ELM-suppression windows. Instead, we
interpolate across these windows using data taken on either side of the windows.

Figure~\ref{fig4} shows the safety-factor in DIII-D discharge \#145380 as a function of time. It can be seen
that $q_{95}\equiv q({\mit\Psi}_N=0.95)$ varies from about 4.1 at the start of the current ramp to about 3.3
at the end. However, the safety-factor at the magnetic axis, $q_0 \equiv q({\mit\Psi}_N=0.0)$, remains approximately
constant at about 1.0. For the sake of practicality,  the {\tt GPEC} and {\tt EPEC} calculations described in this paper ignore $n=3$ rational
surfaces that lie very close to the LCFS: in other words,  surfaces that lie in the region $0.997 < {\mit\Psi}_N\leq 1.000$. The centers of the
first, second, and third ELM-suppression windows correspond to $q_{95}\simeq 3.95$, 3.60, and 3.48, respectively.

Note that all quantities used in our theoretical model are derived from experimental measurements taken in DIII-D discharge \#145380, with the exception of the neutral particle data. 
Unfortunately, no neutral particle data is available for DIII-D discharge \#145380.
Hence, we employ neutral particle data derived from a much earlier discharge (see Sect.~III.D of Ref.~\onlinecite{rf2}).\cite{neutral}
In future, data from the new {\tt LLAMA} (Lyman Alpha Measurement Apparatus) system will be routinely available in DIII-D discharges.\cite{llama}
Combining this data  with modeling from the SOLPS-ITER code\,\cite{solps} ought to generate sufficient information to fully determine the
neutral particle data required by the {\tt EPEC} code.

\subsection{Case 1: $\varpi=\varpi_{\perp\,i}$}
The first case that we shall consider is that in which the natural frequency is determined by the local ion fluid; 
in other words, the natural frequency is specified by Eq.~(\ref{e2}).

The upper panel of Fig.~\ref{fig5} shows the natural frequencies, in the absence of the RMP, of the various  $n=3$ tearing modes 
 that are resonant in the pedestal. It can be seen that the natural frequency of the $m=11$ (brown) tearing mode has its main zero-crossing in the middle
of the first ELM-suppression window. Likewise, the natural frequency of the $m=10$ (magenta) tearing mode has its main zero-crossing in the middle of the
second ELM-suppression window. However, the natural frequency of the $m=9$ (cyan) tearing mode  has its main zero-crossing slightly
prior to the occurrence of the third ELM-suppression window. The lower panel of Fig.~\ref{fig5} shows the natural frequencies, in the presence of the RMP, of the various $n=3$
tearing modes that are resonant in the pedestal. It is apparent that if the magnitude of the natural frequency in the absence of the RMP falls below about
$5\,{\rm krad/s}$ then the associated tearing mode locks to the RMP (i.e., its true natural frequency becomes zero).

The upper panel of Fig.~\ref{fig6} shows the widths of the various vacuum magnetic island chains driven by the applied $n=3$ RMP in the pedestal region as functions of time.
The vacuum island widths are the  expected island widths in the absence of any shielding of driven magnetic reconnection due to
plasma flow; in other words, they are the  widths of the island chains that would be driven if all of the natural frequencies were zero. It can be seen that
the driven magnetic island chains extend over most of the pedestal, implying a very significant degradation in pedestal energy and particle confinement due to the
expected flattening of the density and temperature profiles across the chains. Moreover, the driven island chains at the
bottom of the pedestal overlap, implying that the magnetic field in this region is rendered stochastic. The lower panel
of Fig.~\ref{fig6} shows the actual widths of the various magnetic island chains driven by the applied $n=3$ RMP as functions of time. It is apparent 
that the true island widths are generally much smaller than the vacuum island widths, as a consequence of the shielding of driven magnetic
reconnection due to plasma flow. However, the shielding breaks down at various rational surfaces when the associated natural frequency is zero (see the lower panel of
Fig.~\ref{fig5}). In this situation, the driven island width is similar to the vacuum island width.

Figure~\ref{fig7} shows the widths of the regions of density and electron temperature flattening associated with the $n=3$ magnetic
island chains driven by the RMP in the pedestal region as functions of time. (See Sect.~III.F in Ref.~\onlinecite{rf2}.) (The particle and energy perpendicular diffusivities are assumed to be the
same as the momentum perpendicular diffusivity.) It can be seen that the driven magnetic island chains are generally too narrow
to significantly flatten the density and electron temperature profiles,\cite{flat}  implying that there is actually very little degradation of the pedestal energy and particle confinement
due to the applied RMP. 
However, in those situations in which the shielding of driven
magnetic reconnection due to plasma flow breaks down (i.e., the true natural frequency becomes zero), locked magnetic island chains are generated that are
 wide enough to flatten both the density and the electron temperature profiles.

 Ideally, in order to account for the
ELM-suppression windows in DIII-D discharge \#145380, according to the standard
hypothesis of how RMP-induced ELM-suppression operates, we require the major regions of density and temperature
flattening to correspond to the intersections of the yellow vertical bands and the horizontal dotted lines in Fig.~\ref{fig7}. If this is the case then
the density and temperature profiles are locally flattened at the top of the pedestal (which corresponds to
the horizontal dotted lines) during the ELM-suppression windows (which correspond to the yellow vertical bands). It can be seen
that the region of density and temperature flattening associated with the $m=11$ (brown) tearing mode driven by the RMP lines up fairly well with the
first ELM-suppression window. The region of density and temperature flattening associated with the $m=10$ (magenta) tearing mode 
lines up somewhat less well with the second ELM-suppression window. To be more exact, the region is located at the top of the pedestal,
but occurs slightly too early. However, the 
region of density and temperature flattening associated with the $m=9$ (cyan) tearing mode 
lines up very poorly with the third ELM-suppression window. To be more exact, the region is not really located at the top of the pedestal, and occurs
significantly too early. The ultimate reason for this poor alignment is that the major zero-crossing of the $m=9$ natural
frequency in the upper panel of Fig.~\ref{fig5} does not take place within the third ELM-suppression window (unlike the
major zero-crossings of the $m=11$ and $m=10$ natural frequencies, which do take place within the first and the second
ELM-suppression windows, respectively).

\subsection{Case 2: $\varpi=\varpi_{\perp\,EB}$}
The second case that we shall consider is that in which the natural frequency is determined by the local guiding center fluid; 
in other words, the natural frequency is specified by Eq.~(\ref{e3}).

Figures~\ref{fig8}--\ref{fig10} show analogous data to Figs.~\ref{fig5}--\ref{fig7}, respectively. It can be seen from the
upper panel of Fig.~\ref{fig8} that, in the absence of the RMP,  the natural frequency of the $m=11$ (brown), the $m=10$ (magenta), and the $m=9$ (cyan), tearing modes have their
zero-crossings in the middles
of the first, second, and third, ELM-suppression window, respectively. As before, it is clear from the lower panel of Fig.~\ref{fig8} that these tearing modes lock
to the RMP when the magnitudes of their unperturbed natural frequencies fall below about 5 krad/s. 
Likewise, it is apparent from Fig.~\ref{fig9} that the true island widths driven in the pedestal by the applied
RMP are much smaller than the corresponding vacuum island widths, except when the islands lock to the RMP.
Furthermore, according to Fig.~\ref{fig10}, the regions of density and temperature flattening associated
with the $m=11$ (brown), the $m=10$ (magenta), and the $m=9$ (cyan), tearing modes driven by the RMP line up
very well with the first, second, and third, ELM-suppression windows, respectively. To be more exact, the regions are located at
the top of the pedestal, and also occur at about the correct times. Note that overlapping island chains
are also driven at the bottom of the pedestal. We associate these chains with the density pump-out phenomenon.\cite{hu,rf2} The chains at the bottom of the pedestal are more apparent
in this particular case than in the other two cases considered in this paper because the diamagnetic contributions to the natural frequencies at the bottom of the pedestal (which are absent
in the present case) are comparatively large.

\subsection{Case 3: $\varpi=\varpi_{\perp\,e}$}
The final case that we shall consider is that in which the natural frequency is determined by the local electron fluid; 
in other words, the natural frequency is specified by Eq.~(\ref{e1}).
